Reanalysis Step 5: Sensitivity Analysis with E-values

### 2025-08-20 reanalysis-5-sensitivity.qmd
rework part 5

message("Main Effect of Lapsing on Vision Impairment:")
message("  Odds Ratio: ", round(or_main, 3))
message("  95% CI: (", round(ci_lower_main, 3), ", ", round(ci_upper_main, 3), ")")

@@

# Interpretation
if (evalue_main$est.Evalue > 2) {
  message("\nInterpretation: The E-value of ", round(evalue_main$est.Evalue, 2), 
          " suggests that unmeasured confounding would need to be relatively strong",
          " to explain away the observed association.")
} else if (evalue_main$est.Evalue > 1.5) {
  message("\nInterpretation: The E-value of ", round(evalue_main$est.Evalue, 2),
          " suggests moderate robustness to unmeasured confounding.")
} else {
  message("\nInterpretation: The E-value of ", round(evalue_main$est.Evalue, 2),
          " suggests the finding could be sensitive to unmeasured confounding.")
}
```

@@

message("\nMost robust finding:")
message("  Subgroup: ", most_robust$DR_Severity, " with ", most_robust$Treatment)
message("  E-value: ", most_robust$E_Value_Point)

message("\nLeast robust finding:")
message("  Subgroup: ", least_robust$DR_Severity, " with ", least_robust$Treatment)
message("  E-value: ", least_robust$E_Value_Point)
```

@@

message("Benchmarking E-values against measured confounders in our model:")
message("\nNote: These are approximations based on the associations in our data")

@@

  # Compare to E-values
  max_measured_or <- max(benchmark_table$OR, 1/min(benchmark_table$OR))

  message("\nStrongest measured confounder OR: ", round(max_measured_or, 2))
  message("Minimum E-value across subgroups: ", round(min_evalue, 2))

  if (min_evalue > max_measured_or) {
    message("\nInterpretation: Even the smallest E-value exceeds the strongest",
            " measured confounder, suggesting good robustness.")
  } else {
    message("\nInterpretation: Some E-values are comparable to measured confounders,",
            " suggesting potential sensitivity to unmeasured confounding of similar strength.")
  }
}
```

@@

message("E-values under different outcome prevalence assumptions:\n")

# Get overall outcome prevalence from first imputed dataset
first_data <- readRDS(file.path(reanalysis_data_dir, "imputed_dataset_1.rds"))
outcome_prev <- mean(first_data$outcome_va_vi_binary == 1, na.rm = TRUE)

message("Actual outcome prevalence in data: ", round(outcome_prev * 100, 1), "%")

@@

if (outcome_prev > 0.1) {
  message("\nNote: Since outcome prevalence is ", round(outcome_prev * 100, 1), 
          "%, the 'Common Outcome' assumption may be more appropriate.")
}
```

@@

```{r summary}
# Overall summary
message("=== SENSITIVITY ANALYSIS SUMMARY ===\n")

message("Main Findings:")
message("  - Overall E-value: ", round(evalue_main$est.Evalue, 2))
message("  - Range of subgroup E-values: ", round(min_evalue, 2), " to ", round(max_evalue, 2))

@@

message("\nRobustness Classification:")
message("  - Robust to unmeasured confounding (E > 2): ", robust_subgroups, " subgroups")
message("  - Moderately robust (1.5 < E ≤ 2): ", moderate_subgroups, " subgroups")
message("  - Potentially sensitive (E ≤ 1.5): ", sensitive_subgroups, " subgroups")

# Clinical interpretation
message("\nClinical Interpretation:")
if (min_evalue > 1.5) {
  message("  All subgroup findings show at least moderate robustness to unmeasured confounding.")
} else {
  message("  Some subgroups may be sensitive to unmeasured confounding.")
  message("  Findings should be interpreted with appropriate caution.")
}

# Save E-value results
saveRDS(evalue_table, file.path(reanalysis_data_dir, "evalue_results.rds"))
message("\nE-value results saved to: ", file.path(reanalysis_data_dir, "evalue_results.rds"))
```

>>>>

```

**Main Effect of Lapsing on Vision Impairment:**  
- Odds Ratio: `r round(or_main, 3)`  
- 95% CI: (`r round(ci_lower_main, 3)`, `r round(ci_upper_main, 3)`)

```{r overall-evalue-cont}
#| echo: false

@@

```

**Interpretation:** `r if(evalue_main$est.Evalue > 2) paste0("The E-value of ", round(evalue_main$est.Evalue, 2), " suggests that unmeasured confounding would need to be relatively strong to explain away the observed association.") else if(evalue_main$est.Evalue > 1.5) paste0("The E-value of ", round(evalue_main$est.Evalue, 2), " suggests moderate robustness to unmeasured confounding.") else paste0("The E-value of ", round(evalue_main$est.Evalue, 2), " suggests the finding could be sensitive to unmeasured confounding.")`

@@

```

**Most robust finding:**  
- Subgroup: `r most_robust$DR_Severity` with `r most_robust$Treatment`  
- E-value: `r most_robust$E_Value_Point`

**Least robust finding:**  
- Subgroup: `r least_robust$DR_Severity` with `r least_robust$Treatment`  
- E-value: `r least_robust$E_Value_Point`

@@

```

**Benchmarking E-values against measured confounders in our model:**

*Note: These are approximations based on the associations in our data*

```{r benchmark-cont}
#| echo: false

@@

  # Compare to E-values
  max_measured_or <- max(benchmark_table$OR, 1/min(benchmark_table$OR))
}
```

`r if(exists("max_measured_or")) paste0("**Strongest measured confounder OR:** ", round(max_measured_or, 2), "\n\n**Minimum E-value across subgroups:** ", round(min_evalue, 2), "\n\n**Interpretation:** ", if(min_evalue > max_measured_or) "Even the smallest E-value exceeds the strongest measured confounder, suggesting good robustness." else "Some E-values are comparable to measured confounders, suggesting potential sensitivity to unmeasured confounding of similar strength.") else ""`

```{r benchmark-cleanup}
#| echo: false
#| eval: false
```

@@

```

### E-values under different outcome prevalence assumptions

```{r outcome-calc}
#| echo: false
# Get overall outcome prevalence from first imputed dataset
first_data <- readRDS(file.path(reanalysis_data_dir, "imputed_dataset_1.rds"))
outcome_prev <- mean(first_data$outcome_va_vi_binary == 1, na.rm = TRUE)
```

**Actual outcome prevalence in data:** `r round(outcome_prev * 100, 1)`%

```{r outcome-cont}
#| echo: false

@@

```

`r if(outcome_prev > 0.1) paste0("*Note: Since outcome prevalence is ", round(outcome_prev * 100, 1), "%, the 'Common Outcome' assumption may be more appropriate.*") else ""`

@@

```{r summary}
# Calculate summary statistics

@@

# Save E-value results
saveRDS(evalue_table, file.path(reanalysis_data_dir, "evalue_results.rds"))
```

### SENSITIVITY ANALYSIS SUMMARY

**Main Findings:**  
- Overall E-value: `r round(evalue_main$est.Evalue, 2)`  
- Range of subgroup E-values: `r round(min_evalue, 2)` to `r round(max_evalue, 2)`

**Robustness Classification:**  
- Robust to unmeasured confounding (E > 2): `r robust_subgroups` subgroups  
- Moderately robust (1.5 < E ≤ 2): `r moderate_subgroups` subgroups  
- Potentially sensitive (E ≤ 1.5): `r sensitive_subgroups` subgroups

**Clinical Interpretation:**  
`r if(min_evalue > 1.5) "All subgroup findings show at least moderate robustness to unmeasured confounding." else "Some subgroups may be sensitive to unmeasured confounding. Findings should be interpreted with appropriate caution."`

**E-value results saved to:** `r file.path(reanalysis_data_dir, "evalue_results.rds")`
